Large-scale harmonic patterns clustering

### 2010-03-20 Paper/Paper.tex
stuff added back from draft

\section{Previous Work}\label{sec:prevwork}
Casey and Slaney \cite{Casey2006,Casey2007,Casey2008}

@@

\subsection{EchoNest API}

@@

\begin{algorithm}
%\caption{Pseudocode of Vector Quantization}
\begin{algorithmic}
\STATE$l$ learning rate
\STATE$\{P_n\}$ set of patches
\STATE$\{C_k\}$ codebook of $K$ codes
%\STATE $m \leftarrow b$
\REQUIRE $0 < l \leq 1$
\FOR{$nIters$}
\FOR{$p \in \{P_n\}$}
\STATE$c \leftarrow min_{c \in C_k} dist(p,c)$
\STATE$c \leftarrow c + (p - c) * l$
\ENDFOR
\ENDFOR
\RETURN $\{C_k\}$
\caption{{Pseudocode of Online Vector Quantization. Note that we can replace
the number of iteration by a threshold on the distortion over some test set.}
\label{algo:vq}}
\end{algorithmic}
\end{algorithm}

@@

\section{Conclusion and Future Work}

>>>>

\subsection{Definitions}
A pattern, or a patch, is a \textit{Shingle} taken from an actual song.
A code is similar to a pattern, but is the result of the clustering
algorithm and most likely does not come from an actual song.
Distortion is our measure of distance between a pattern and a code
(or between two patterns, see equation \ref{eq:dist} in Section 
\ref{sec:experiments}).

\section{Previous Work}\label{sec:prevwork}
Our work can be seen as an extension of the \textit{Shingles} idea
described by Casey and Slaney \cite{Casey2006,Casey2007,Casey2008}.
The two main differences are the clustering framework and the use of
beats and bars to select the position of the \textit{Shingles}.

Casey and Slaney \cite{Casey2006} use LSH \cite{Datar2004} to perform
fast approximate nearest neighbor search. Their implementation is E$^2$LSH
\cite{E2LSH}. This package can efficiently handle about $50$K or $100$K
vectors of size $200$. Even with a better implementation, the LSH
model still has some scaling problem. A large portion of the data
must be known to decide how many projections to use, and what guarantee
of finding the nearest neighbor it can provides. Also, we could fix the
number of projections and consider as clusters all the elements that
fall into the same bin. However, this does not give us a centroid, a
typical pattern representing that cluster.

@@

\subsection{EchoNest API}
We use the Echo Nest analyze API \cite{EchoNest} throughout this work.
The API gives us a chroma vector (length $12$) for every music event, 
or ``segment'', for any song we upload to their platform. 
We also get an estimate of where the beats and the bars are. Beats span over
multiple segments, and bars span over multiple beats. 
We transform the chroma vectors per segment into per beat vectors using a 
simple average. We can then stack beat vectors for a number of bars 
(usually 1 or 2). 
We get a fixed-size patch by resampling the number of chroma vectors. Typically,
there are $4$ beats per bar. Thus an appropriate pattern size for one bar
is $12$x$4$ or $12$x$8$.

Note that we do not claim that any of these informations (segments, beats, bars)
are perfectly accurate (e.g. in \cite{Barrington2009a} authors obtain a better
song segmentation than Echonest). Practise showed us that they are reasonable, 
and the size of the data set should make up for the imperfections or noise.
We also believe that patches sizes based on a number of beats or bars are more
meaningful than an arbitrary time length. More on this in the experiments
(Section \ref{sec:experiments}).

\begin{figure}[htb]
\begin{center}
\includegraphics[width=.8\columnwidth]{code}
\end{center}
\caption{\small{A typical code from a codebook of size $200$. 
Patterns represented
$2$ bars and the pattern length was set to $16$.
BLACK AND WHITE VERSION COMING.
}}
\label{fig:code}
\end{figure}

@@

\begin{algorithm}
%\caption{Pseudocode of Vector Quantization}
\begin{algorithmic}
\STATE$l$ learning rate
\STATE$\{P_n\}$ set of patches
\STATE$\{C_k\}$ codebook of $K$ codes
%\STATE $m \leftarrow b$
\REQUIRE $0 < l \leq 1$
\FOR{$nIters$}
\FOR{$p \in \{P_n\}$}
\STATE$c \leftarrow min_{c \in C_k} dist(p,c)$
\STATE$c \leftarrow c + (p - c) * l$
\ENDFOR
\ENDFOR
\RETURN $\{C_k\}$
\caption{\small{Pseudocode of Online Vector Quantization. Note that we can 
replace the number of iteration by a threshold on the distortion over some 
test set.}
\label{algo:vq}}
\end{algorithmic}
\end{algorithm}

@@

\section{Conclusion and Future Work}
We presented a practical method to perform large-scale clustering of
harmonic patterns. We assessed the basic properties of the method through
experiments on a large collection of music. We suggested many ways
to interpret the data through visualization (LLE) and further encoding
(NMF). We discussed the possibility to move to even larger scales
and we provided our source code\footnote{Code not yet released to preserve
submission's anonymity}.

As for enhancements, we specifically did not compress our features using
gaussian mixtures or other generative model. That being said, these methods
could be use to develop better distance measures between \textit{Shingles}.
The use of the euclidean distance is arbitrary. Summarizing patches
with gaussians, and then comparing the distance between those gaussians,
could reduce the influence of the noise in the distance measure.

Moving on to larger scales, we could imagine merging codes together and
splitting some on two if the clusters become too small or too large.We could
also cluster the codebook itself, in a similar fashion as hierarchical
gaussian mixtures \cite{Vasconcelos2001}.
